---
title: "Aids Drug"
author: "Ever De Leon"
date: "03/11/2024"
output: html_document
---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
triUmp <-read.csv(file="triUmpsal.csv",header=TRUE)
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?.  If you skimp in this section expect to be marked down in your grade. Paint a picture of the project a couple of paragraphs is expected!-->

A new preventative aides drug (triUmpsal) is under test in clinical trials. Consider only the DrugUsers, does this drug seem effective?  Hint - use dplyr to obtain that subpopulation with the filter command.

$H_0$: The new preventative aides drug does not seem effective in helping those who have been diagnosed with aids.

$H_A$: The new preventative aides drug weems to be effective in helping those who have contracted aids. 

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method. Discuss what data manipulations may be needed for example subsetting or other binning procedures and what sort of graphical displays and inferential tools will be used!-->


Variables we use are:

Treatment - categorical.  
DrugUser - categorical.   
Infected - categorical.

Infected depends on treatment among users.

We get just the drug users from the population.

```{r}
drugUserData = select(triUmp, treatment, DrugUser, infected)
drugUserData = filter(drugUserData, DrugUser== "yes")
```

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question. Bottom line is to explain the features of the graph that are relevant to conforming or departing to what is expected by your null hypothesis !-->

```{r}
barchartGC(~infected + treatment,data=drugUserData)
barchartGC(~infected + treatment,data=drugUserData,  type="percent")
```

The percent bar chart shows that among those that have been given the placebo, that percentage has risen whereas those who have been getting the drug shows a lower percentage showing that the drug is working.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
table1 <- xtabs(~infected + treatment, data=drugUserData)
rowPerc(table1)
colPerc(table1)
```

In the top graph, if your infected with aids then 24.29% were getting the drug and 75.71% were getting the placebo. If the drug was not effective then the numbers would be at closer to 50% and 50%. Since they are not, then the drug may be working.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

In these test runs, we see that the p-value is 0.0000000423. This means that the preventative aids drug seems effective.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

We are trying to see if this new preventative aids drug is effective. The p-value is 0.0000000423, which is less than 0.05 so we reject the $H_0$. We conclude that the drug is effective. When reading the bar chart, it looks as though the preventative new aids drug may be effective. In the odds ratio, when usung the drug, the odds are .35 to 1 that you will be cured from aids when using the treatment.